---
title: "week 4 data analysis"
author: "keyla"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Overview

This report represents preliminary data analysis for the week 3 example data. Basically this will report basic descriptive statistics as well as report tests of the following predictions:

1. People will feel more guilt when:
  + 1.1 They cause the shock (vs. their partner)
  + 1.2 Their partner was morally good (vs.bad)
  + 1.3 The shock was stronger (high > medium > low)
2. Women will report more guilt than men
3. Agreeable people will report more guilt
4. Neurotic people will report more guilt

and

5. People will be more generous when:
  + 5.1 They caused the shock (vs. their partner)
  + 5.2 The partner was morally good (vs.bad)
  + 5.3 The shock was stronger (high > medium > low)
6. Women will be more generous than men
7. Agreeable people will be more generous
8. Neurotic people will be more generous


```{r load}
data<-read.csv("week 3/processed data/Week 3 DATA PROCESSED.csv")
library(reshape2)
#Recode variables:
data$PIN<-as.factor(data$PIN)
data$sex<-as.factor(data$sex)
data$shock<-as.factor(data$shock)
data$shockCause<-as.factor(data$shockCause)
data$pMoral<-as.factor(data$pMoral)
```



## Descriptives

Descriptive statistics for guilt:
```{r descGuilt}

c(summary(data$guilt),"SD"=sd(data$guilt,na.rm = T))

```


Descriptive statistics for generosity:
```{r descGenerosity}

#Generosity before the shock:
c(summary(data$preShare),"SD"=sd(data$preShare,na.rm = T))


#Generosity after the shock:
c(summary(data$postShare),"SD"=sd(data$postShare,na.rm = T))

```


Descriptive statistics for personality:
```{r descPers}

#Descriptives for agreeableness

c(summary(data$agree),"SD"=sd(data$agree,na.rm = T))


#Descriptives for neuroticism

c(summary(data$neuro),"SD"=sd(data$neuro,na.rm = T))
```



## Guilt Analyses

### Prediction 1.1

```{r p1.1}
#Run the t-test for prediction 1.1
p1.1<-t.test(data$guilt~data$shockCause)

#Output the result
p1.1

```

### Prediction 1.2
Prediction 1.2 was supported. People felt more guilty when their partner was morally good.
```{r p1.2}
p1.2<-t.test(data$guilt~data$pMoral)
p1.2

```

### Prediction 1.3
```{r p1.3}
#run the one way ANOVA for prediction 1.3
p1.3<-lm(guilt~shock,data = data)

#output the result
anova(p1.3)

#output group means
tapply(data$guilt,data$shock,function(x)mean(x,na.rm=T))
```

### Prediction 2
```{r p2}
#run t test for prediction 2
p2<-t.test(data$guilt~data$sex)
p2
```

### Prediction 3
```{r p3}
p3<-cor.test(data$guilt,data$agree)
p3
```
###Prediction 4
```{r p.4}
p.4<-cor.test(data$guilt,data$neuro)
p.4
```

## Generosity Analysis
```{r melt}
dataMelt<-melt(data,
               measure.vars = c("preShare","postShare"),
               variable.name = c("prePost"),
               value.name = c("share")
               )
```
### Prediction 5.1
```{r p5.1}
p5.1<-aov(share~prePost*shockCause+Error(PIN),data=dataMelt)
summary(p5.1)

tapply(dataMelt$share,dataMelt[,c(24,21)],function(x)mean(x,na.rm=T))
```

